---
title: "Google Summer of Code Contributions"
author: "Lorenz Walthert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Google Summer of Code Contributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette aims to summarize the contributions made to styler through Google
Summer of Code 2017.
```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
knitr::knit_engines$set(list(
  styler = function(options) {
    options$comment <- ""
    knitr::engine_output(
      options,
      
      c("# Before", options$code),
      c("# After", styler::style_text(options$code))
    )
  }
))

```

# styler in practice
As a warm-up, we want to give a few examples of how styler styles code. It's
possible to use different levels of invasiveness, as described in the help
file for only style guide implemented so far, which is the [tidyverse style guide](http://style.tidyverse.org/index.html). The style guide in use is 
passed to the styling function (i.e `style_text()` and friends) via 
`style` argument, which defaults to `tidyverse_style`.
Apart from this argument, there are further customization options. For example, 
we can limit ourselves to styling just spacing information, which we can 
indicate with the `scope` argument.
```{r}
library("styler")
library("magrittr")
style_text("a=3; 2", scope = "spaces") %>% 
  cat()
```

Or, on the other extreme of the scale, styling spaces, indention, line breaks 
and tokens:
```{r}
style_text("a=3; 2", scope = "tokens") %>%
  cat(sep = "\n")
```

Another option that is helpful to determine the level of invasiveness is
`strict`. If set to `TRUE`, spaces and line breaks after or before tokens are
set to either zero or one. However, in some situations that might be undesirable,
as the following example shows:
```{r}
style_text(
  "data_frame(
     small  = 2 ,
     medium = 4,#comment without space
     large  = 6
   )", strict = FALSE) %>%
  cat(sep = "\n")
```

We prefer to keep the equal sign after "small", "medium" and large "alligned",
so we set `strict = FALSE` to set spacing to *at least* one around `=`.

I hope these examples showed you some of the options available in styler. Let
us for now focus on a configuration with `strict = TRUE` and `scope = "tokens"` 
and therefore only show code before and after styling. 

styler can distinguish unary operators and other math tokens:

```{styler}
1++1-1-1/2
```

It can also format complicated expressions that involve line breaking and 
indention based on both brace expressions and operators:
```{styler}
if (x >3) {stop("this is an error")} else {
c(there_are_fairly_long,
1 / 33 * 
2 * long_long_variable_names)%>% k(

) }
```

Lines are broken after `(` if a function calls spans over multiple lines:

```{styler}
do_a_long_and_complicated_fun_cal("which", has, way, to, 
                              "and longer then lorem ipsum in its full length"
                              )
```

Styler converges `=` assignment, Single quotes if necessary, adds braces to
function calls in pipes:
```{styler}
one= 'one string'
two= "one string in a 'string'"
a %>%
  b %>%
  c

```

Function declarations are indented if multi-line
```{styler}
my_fun <- function(x, 
y, 
z) {
  just(z)
}
```

Styler can also deal with tidyeval syntax
```{styler}
mtcars %>%
  group_by( ! !!my_vars)
```

# Contributions to styler during GSOC
## High-level analysis
The author, with help of his mentors,

* opened over [60](https://github.com/krlmlr/styler/issues?utf8=✓&q=is%3Aopen%20is%3Aissue%20author%3Alorenzwalthert%20created%3A%3C2017-08-31%20%20created%3A%3E2017-06-01%20) issues and closed the majority of them with his own pull requests.
* created more than [60](https://github.com/krlmlr/styler/pulls?q=is%3Apr+author%3Alorenzwalthert+created%3A%3C2017-08-31++created%3A%3E2017-06-01) pull requests to
  master of which the vast majority was merged.

The activity during GSOC can be grouped as follows:

- **Replicating the outcomes that were achieved with flat styling with nested styling.** 
  Required implementation of the visitor as well as different 
  serialization method.
- **Adding functionality that was only possible with the nested styling** such 
  as indention based on operators (`+`, `%>%` etc.), braces `{` and `(`. 
  the use of raw indention, allowing for token-dependent re-indention.
- **Infrastructure changes** such as serialized testing, improve performance, 
  create mutli-line attribute, code re factoring. Partially
  flattening out the nested parse table.
- **Extending the functionality that was not there with the flat approach.** 
  Examples include styling multiple expressions, token replacement (e.g. `<-` for 
  `=`), line break rules (invasive and non-invasive), token insertion (`()` in 
  function calls within pipes), RStudio Addins, styling of comments, styling of 
  a single file, public API redesign, spacing around more operators (colons, 
  commas, infix operators, bang-bang).

The list above is not comprehensive and covers only the large contributions.
Always along with coding, documentation and tests were provided. With regard
to the initial [proposal](https://docs.google.com/document/d/1V7ZWOVjEqM6-vc7WVxePUOHHc48HJwBKh8gAwIFALJ4/edit), we can say that steps one to four (increase documentation and 
testing of code, perform all operations on with a nested structure, and fully 
support the tidyverse style guide) were completed to a large extent and a few 
points under step five have been implemented too. The remaining points under 
step five as well as the open issues of the repository indicate the direction of
further development of the package.

The tidyverse style guide does not cover all formatting cases, so in particular 
for function calls, there are various cases uncovered and a corresponding 
[issue](https://github.com/krlmlr/styler/issues/125) was opened in the styler 
repository.
Also, it was not possible in this project to implement the rule that forces
lines to be less than 80 characters long. We leave this to [lintr](https://github.com/jimhester/lintr). Known remaining problems are 
well-documented on the [issue page](https://github.com/krlmlr/styler/issues) of
the main repository on GitHub.
It is nice to see that `styler` already got some traction from the community.
The maintainer of the package [reprex](https://github.com/tidyverse/reprex) 
plans to [add styler as a dependency](https://github.com/tidyverse/reprex/issues/94).
As soon as one of its dependencies is accepted to CRAN, styler will also be 
submitted.

# Low-level analysis

We can use the package `gitsum` to perform an analysis of the git repository.
```{r, message = FALSE}
library("gitsum")
library("tidyverse")
log <- git_log_detailed() %>%
  filter(date > as.Date("2017-06-01"), date < as.Date("2017-08-31"))
```

We can see that the vast majority of the commits in this period comes from the 
student and a few from the main mentor.
```{r}
log %>%
  group_by(author_name) %>%
  count()
```


We can also look at how the number of lines in the different files evolved:
```{r}
lines <- log %>%
  add_line_history()

r_files <- grep("^R/", lines$changed_file, value = TRUE)

to_plot <- lines %>%
  filter(changed_file %in% r_files) %>%
  group_by(changed_file) %>%
  mutate(has_changed = n() > 1) %>%
  filter(has_changed)
  
ggplot(to_plot, aes(x = date, y = current_lines)) + 
  geom_step() + 
  scale_y_continuous(name = "Number of Lines", limits = c(0, NA)) + 
  facet_wrap(~changed_file, scales = "free_y")
```

The total number of insertions and deletions by the author is
```{r}
log %>%
  filter(author_name == "Lorenz Walthert") %>%
  summarize(sum(total_insertions) + sum(total_deletions)) %>%
  pull()
```

Which still contains many insertions and deletions in some files that were
generated as a bi-product, so let's focus on the code in the `R` directory.

```{r}
log %>%
  unnest() %>%
  filter(dirname(changed_file) == "R") %>%
  summarize(sum(insertions) + sum(deletions)) %>%
  pull()
```

The core code with .R files only consists of 

```{r}
log %>%
  unnest() %>%
  filter(grepl("\\.R$", basename(changed_file))) %>%
  summarize(sum(insertions) - sum(deletions)) %>%
  pull() %>%
  cat("lines")
  
```

# Final remarks

First and foremost, I want to thank my mentors, in particular Kirill Mueller, for his patience and time he put into that project. I really appreciate that and I think apart from the code submitted, I benefited so much from him in terms of how to code, what considerations to make and how to deal with pitfalls and unexpected problems along the way.

In addition, I want to thank Google for their contributions to open source that made this project possible as well as all people involved with organizing GSOC from the R community. It was an amazing experience.

